\begin{thebibliography}{99}

\bibitem{AGK}
D.~Armbruster, J.~Guckenheimer, and S.~Kim
\newblock Chaotic dynamics in systems with square symmetry.
\newblock Physics Letters A, 140, 1989.

\bibitem{AVK}
A.~Andronov, A. Vitt, and S. Khaiken.
\newblock {\it Theory of Oscillations}.
\newblock New York, Pergamon, 1966.

\bibitem{arnold}
V.~I.~Arnold.
\newblock {\em Mathematical Methods of Classical Mechanics}, 2nd Ed.
\newblock Springer-Verlag, New York, 1989.

\bibitem{BGKM}
C.~Baesens, J.~Guckenheimer, S.~Kim, and R.~Mackay.
\newblock Three coupled oscillators: mode-locking, global
bifurcations, and toroidal chaos.
\newblock Physica D 49, 1991.

\bibitem{BGKM:IMA}
C.~Baesens, J.~Guckenheimer, S.~Kim, and R.~Mackay.
\newblock Simple resonance regions of torus diffeomorphisms.
\newblock IMA preprint series \#656, 1990.

\bibitem{BC} 
M. Benedicks and L. Carleson.
\newblock The dynamics of the H\'{e}non map.
\newblock Annals of Mathematics, 133, 1991.

\bibitem{chirikov}
B. V. Chirikov.
\newblock A universal instability of many-dimensional oscillator
systems.
\newblock Physics Reports, 52, 1979.

\bibitem{dangle1}
G.~Danglemayr and J.~Guckenheimer
\newblock On a Four Parameter Family of Planar Vector Fields
\newblock Archive for Rational Mechanics and Analysis, Vol. 97, No. 4, 1987, pps. 321-352.

\bibitem{golubitsky1}
P.~Chossat and M.~Golubitsky. 
\newblock Symmetry increasing bifurcation of chaotic attractors. 
\newblock Physica D 32 (1988).

\bibitem{Deng:constructing}
B.~Deng.
\newblock Constructing homoclinic orbits and chaos.
\newblock Preprint, April 1991.

\bibitem{Devaney}
R.~Devaney.
\newblock {\em An Introduction to Chaotic Dynamical Systems}.
\newblock  Addison-Wesley, Redwood City, CA, 1989.

\bibitem{epsteinmarder}
I. R. Epstein and E.~Marder.
\newblock Multiple modes of a conditional neural oscillator
\newblock Biol. Cybern. 63, 1990.

\bibitem{golubitsky2}
M. Field and M. Golubitsky.
\newblock Symmetric chaos.  
\newblock Computers in Physics.   Sep./Oct. 1990.

\bibitem{golubitsky3} 
M.~Golubitsky and D. G.~Schaeffer.
\newblock {\it Singularities and groups in bifurcation theory}.
\newblock Applied Mathematical Sciences Series, Vol. 51, Springer-Verlag,  1985.

\bibitem{gucken1} J.~Guckenheimer
\newblock Multiple bifurcation problems for chemical reactors.
\newblock Physica D 20, 1986.

\bibitem{gucken2}
J.~Guckenheimer
\newblock Dynamics of the Van der Pol Equation
\newblock IEEE Transactions on Circuits and Systems, Vol. CAS-27, No. 11, 1980, pps. 983-989.

\bibitem{GGHW}
J. Guckenheimer, S. Gueron and R. Harris-Warrick.
\newblock The dynamics of a conditionally bursting neuron.
\newblock Preprint, 1992.

\bibitem{GH}
J.~Guckenheimer and P.~J. Holmes.
\newblock {\em Nonlinear Oscillations, Dynamical Systems, and Bifurcations of
  Vector Fields}.
\newblock Springer-Verlag, New York, 1983.

\bibitem{GuckenMahalov}
J.~Guckenheimer and A.~Mahalov.
\newblock Phys. Rev. Lett. 68, 1992.

\bibitem{GuckenheimerWorfolk:instant}
J.~Guckenheimer and P.~Worfolk.
\newblock Instant chaos.
\newblock Nonlinearity, to appear, 1992.

\bibitem{Henon:mapping}
M.~H\'{e}non.
\newblock A two-dimensional mapping with a strange attractor.
\newblock {\em Comm. in Mathematical Physics}, 50:69--77, 1976.

\bibitem{hh1}
A.L.~Hodgkin and A.F. Huxley
\newblock {A Quantitative Description of Membrane Current and its Applications
  to Conduction and Excitation in Nerve}. 
\newblock Journal of Physiology, Vol. 117, 1952, pps. 500-544.

\bibitem{kubicek1} 
M.~Kubi\u{c}ek
\newblock Evaluation of complex bifurcation points.
\newblock SIAM Journal of Applied Mathematics, Vol. 38, No. 1, 1980.

\bibitem{lorenz1}
E.N.~Lorenz
\newblock Deterministic Non-Periodic Flows
\newblock Journal of Atmospheric Science, Vol. 20, 1963, pps. 130-141.

\bibitem{MoraViana:abundance}
L.~Mora and M.~Viana.
\newblock Abundance of strange attractors.
\newblock preprint, 1991.

\bibitem{rinzel1} 
R.J.~Rinzel and Y.S. Lee
\newblock Dissection of a model for neuronal parabolic bursting.
Journal of Mathematical Biology, Vol. 25, 1987.

\bibitem{rinzel2}
J.~Rinzel and R.N. Miller
\newblock {Numerical calculation of stable and unstable periodic solutions 
	  to the Hodgkin-Huxley equations}.
\newblock Mathematical Biosciences, Vol. 49, 1980, pps. 27-59.

\bibitem{salamsastry} 
F.~Salam and S.~Sastry.
\newblock Dynamics of the forced Josephson junction circuit: the
regions of chaos.
\newblock IEEE Transactions on Circuits and Systems, CAS-32, 8, 1985.

\bibitem{sparrow1}
C.~Sparrow
\newblock {\em The Lorenz Equations: Bifurcations, Chaos, and Strange Attractors}.
\newblock Springer-Verlag Series in Applied Mathematical Sciences, Vol. 41, 1982.

\bibitem{takens1}
F.~Takens
\newblock Forced Oscillations and Bifurcations
\newblock {Communications of the Mathematical Institute, Rijkuniversiteit Utrecht,
           Vol. 3, 1974, pps. 1-59.}

\bibitem{uppal1} 
A.~Uppal, W. H.~Ray, and A.~B.~Poore.
\newblock On the dynamic behavior of continuous stirred tank reactors.
\newblock Chemical Engineering Science, Vol. 29, 1974.

\bibitem{vdp1}
B.~van der Pol
\newblock {Forced Oscillations in a Circuit with Nonlinear Resistance
(Receptance with Reactive Triode)}.
\newblock London, Edinburgh and Dublin Phil. Mag., Vol. 3, 1927, pps. 65-80.

\end{thebibliography}
