\documentstyle[fullpage]{article}

\begin{document}

\title{Weekly Status Report}

\author{Group 7 --- Project Planning Tool\\
Project Manager: Brian Gaubert\\
Project Coordinator: Mike Marlow\\
Documentation Manager: Mark M. Lacey\\
Configuration Manager: Richard Malingkas}

\date{Due October 12, 1992}

\maketitle

%
% This section is used to list the accomplishments of the week.
% Use the \item construct to list each item.  Try to keep the
% Descriptions for each down to one or two sentences
%
% Limit your accomplishments to items specific to the progress of the project.
%
\section{Planned Accomplishments}
\begin{itemize}
	\item Had a group meeting last Sunday to discuss the different
parts of the requirement specifications. We split up the tasks among
the people in the group and discussed each section as it pertains to
the group project as a whole. (1 hr)
	\item The group met on Wednesday to discuss the different
parts of the requirement specifications and how it will be put
together. (3 hrs)
	\item Thursday, the group integrated each part of the
requirement specification. (14 hrs)
	\item Friday, The group proof-read the document and finished
up the minor details. (3 hrs)

\end{itemize}

%
% This section is used to list the unscheduled accomplishments of the week.
% Use the \item construct to list each item.  Try to keep the
% Descriptions for each down to one or two sentences
%
% Limit your accomplishments to items specific to the progress of the project.
%
\section{Other Accomplishments}
\begin{itemize}
	\item Modified the user interface prototype.
\end{itemize}

%
% This section is used to list the following weeks plan
% Use the \item construct to list each item.  Try to keep the
% Descriptions for each down to one or two sentences
%
% Limit your plans to items specific to the progress of the project.
%
\section{Next Weeks Plan}
\begin{itemize}
	\item Met with the group Sunday afternoon. We discussed and
partitioned the sections of the Preliminary Design among the members
of the group. These sections must be done Wednesday for integration.
\end{itemize}

%
% This section is used to list any issues that were raised during
% the week that are of special interest.  Also use this to voice
% issues that the TA or Instructor must resolve.
% Use the \item construct to list each item.  Try to keep the
% Descriptions for each down to one or two sentences
%
% Limit your issues to items specific to the progress of the project.
%
\section{Issues}
\begin{itemize}
	\item The time required for this project, documentation etc.,
is well over the amount of time involved for a normal four credit class.
\end{itemize}

\end{document}

