\documentstyle[fullpage]{article}

\begin{document}

\title{Weekly Status Report}

\author{Group 7 --- Project Planning Tool\\
Project Coordinator: Brian Gaubert\\
Project Manager: Mike Marlow\\
Documentation Manager: Mark M. Lacey\\
Configuration Manager: Richard Malingkas}

\date{Due October 19, 1992}

\maketitle

%
% This section is used to list the accomplishments of the week.
% Use the \item construct to list each item.  Try to keep the
% Descriptions for each down to one or two sentences
%
% Limit your accomplishments to items specific to the progress of the project.
%
\section{Planned Accomplishments}
\begin{itemize}
	\item The group met on October 12 to discuss the preliminary
 design document, and assign individual tasks regarding it. (1.5 hrs)
	\item Preliminary design document sections were revised and integrated into 1 document (5 hrs)
	\item The group met on October 15 to proofread, comment, and offer
necessary revisions on final preliminary design document. (0.5 hrs)
	\item The group met on October 18 to discuss the detailed  
design document, and assign individual tasks regarding it.  We also discussed
how xplan's  user manual will be written and formatted. (1 hr )

\end{itemize}

%
% This section is used to list the unscheduled accomplishments of the week.
% Use the \item construct to list each item.  Try to keep the
% Descriptions for each down to one or two sentences
%
% Limit your accomplishments to items specific to the progress of the project.
%
\section{Other Accomplishments}
\begin{itemize}
	\item None.
\end{itemize}

%
% This section is used to list the following weeks plan
% Use the \item construct to list each item.  Try to keep the
% Descriptions for each down to one or two sentences
%
% Limit your plans to items specific to the progress of the project.
%
\section{Next Weeks Plan}
\begin{itemize}
	\item The group will meet on October 21 to discuss the detailed
 design document.  We will integrate new work into the preliminary
design document's ideas, discuss necessary revisions, work on items such as structure
charts, and program design language descriptions of our software.  Implementation
methods, data structures, algorithms, and a requirements dictionary will be developed.
	\item The group will meet on October 25 to finalize changes after proof reading
the detailed design document.
	\item Discuss and develop the xplan user manual
\end{itemize}

%
% This section is used to list any issues that were raised during
% the week that are of special interest.  Also use this to voice
% issues that the TA or Instructor must resolve.
% Use the \item construct to list each item.  Try to keep the
% Descriptions for each down to one or two sentences
%
% Limit your issues to items specific to the progress of the project.
%
\section{Issues}
\begin{itemize}
\item None.
\end{itemize}

\end{document}


